Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-04-18 CAV2018/paper.tex
Floating point benchmark runner fixes
Updated benchmark references in reviewers questions

%\documentclass[sigconf]{acmart}
%\documentclass[sigconf]{llncs}
\documentclass[a4paper,UKenglish]{lipics-v2018}

\usepackage{microtype}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amssymb}
\usepackage{amsmath}
%\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{framed}
\usepackage{listings}
\usepackage{algorithm,algpseudocode}
\usepackage{epstopdf}
\usepackage{pifont}
\usepackage{todonotes}
\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usetikzlibrary{patterns}
\usepackage{url}
\usepackage{mathtools}
\usepackage{cite}

@@

%\email{alessandro.abate@cs.ox.ac.uk}
%\email{dario.cattaruzza@cs.ox.ac.uk}
%\email{cristina.david@cs.ox.ac.uk}
%\email{kroening@cs.ox.ac.uk}

@@

%\begin{keyword}                   
%Digital Control; A/D converters; Control System Synthesis; Safety Analysis; Quantization Errors; Observer-based Controller. Sampling.
%\end{keyword}

@@

\begin{theorem}
%
The error introduced by a delay $t_d$ in the control input $\vec{u}_k$ updated
by a discrete feedback controller for a continuous LTI system can be modeled as
a non-deterministic disturbance signal ($\nu_t$ in
Figure~\ref{fig:observersystem}) upper bounded by
%
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, $t_s=T_s-t_d$,
and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm, 
which we extend to sets as $\norm{X}_1 = \sup \{\norm{\vec{x}}_1, \vec{x} \in X\}$.  
For the proof see Appendix~\ref{sec:proofs}
\end{theorem}
%
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
This result allows to over-approximate the effects of the delay by introducing a disturbance signal $\nu_t=[-q_t, q_t]$ at the output.
 %and use the dynamics of the system without delay.  
 This noise will be small when the delay is much smaller than the sampling time ($t_s\approx T_s$ and $\mat{B}_{t_d} \approx 0$).

@@

A safety specification leads to a requirement on the states of the model. 
Consequently, the controller must ensure that the state never violates the
requirement.  In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}^n: = & \forall k \in \{0, \ldots , n\},
  \mat{G}\vec{x}_k \leq \vec{h} \text{ and } \underline{y} \leq y_k \leq \overline{y}\nonumber\\
\phi_\mathit{safety}: = & \phi_\mathit{safety}^\infty
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
Note that FWL effects are not explicitly captured here
as they are implicit from Equation \eqref{eq:observer_LTI_FWL_cf}.

@@

We must point out that the optimizer tool uses infinite precision for the
plant, and may therefore find solutions that the synthesizer cannot, due to
the use of limited precision.  If we cannot optimize the candidate we signal
the {\sc Abstract} phase to use the updated $\phi$ and request {\sc
Synthesize} to produce a new candidate.  Otherwise the optimized candidate
solution is given back to {\sc Verify Synthesis}.

@@

The benchmark suite employed to evaluate the proposed technique is composed
by seven models extracted from the control literature~\cite{acrobot, cstr,
CHEN1979389, KOKOTOVIC198023, gajic2008optimal, Franklin15, maglev,
converters}.  In~order to obtain statistically relevant results, we have
taken these models and modified each possible combination of their
parameters by a 10\% tolerance, thus creating seven hundred benchmarks from
this base.

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{order} & \multicolumn{1}{|c|}{No.} & \multicolumn{4}{|c|}{CEGIS} & \multicolumn{4}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 20& 1.1\,s &   3.8\,s & 2 & 58\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 15 & 7.1\,s  &   18.0\,s & 9 & 90\\
3  & Inverted Pendulum & 4 & 100 &  277\,s& t.o. & 82& 0 & 20.0\,s  &  t.o. & 55 & 20\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 42 & 3.0\,s  &  9.5\,s & 5 & 95 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 27 & 71 &0.4\,s  &   4.9\,s & 6 & 92\\
6  & Suspension        & 4 & 100 &15.0\,s & 177.2\,s & 30 & 70 & 14.8\,s&166.9\,s  & 26 & 73\\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11 & 80 &3.1\,s  &   16.8\,s & 8 & 82\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results.  Minimum, and Median times for standard CEGIS vs
CEGIS-OR on multiple benchmarks.  No.  indicates Number of benchmarks,
t.o.~indicates timeouts (in the median case, it indicates more than half of
the benchmarks timed out).  The order denotes the number of variables of a
model.}
%
\end{table}

@@

In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS procedure.  For the majority
of our benchmarks, we observe that our approach is significantly faster than
standard CEGIS.  We further note that the number of times standard CEGIS
times out is much larger than CEGIS-OR which results in more found
solutions.

>>>>

%\documentclass[sigconf]{acmart}
%\documentclass[sigconf]{llncs}
\documentclass[a4paper,UKenglish]{lipics-v2018}
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amssymb}
\usepackage{amsmath}
%\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{framed}
\usepackage{listings}
\usepackage{algorithm,algpseudocode}
\usepackage{epstopdf}
\usepackage{pifont}
\usepackage{todonotes}
\usetikzlibrary{positioning, automata, shapes.arrows, calc, shapes, arrows}
\usetikzlibrary{patterns}
\usepackage{url}
\usepackage{mathtools}
\usepackage{cite}

@@

%\email{alessandro.abate@cs.ox.ac.uk}
%\email{dario.cattaruzza@cs.ox.ac.uk}
%\email{cristina.david@cs.ox.ac.uk}
%\author{Daniel Kroening}\email{kroening@cs.ox.ac.uk}

@@

%\begin{keyword}        
\keywords {Digital Control; A/D converters; Control System Synthesis; Safety Analysis; Quantization Errors; Observer-based Controller. Sampling.}
%\end{keyword}

@@

\begin{theorem}
%
The error introduced by a delay $t_d$ in the control input $\vec{u}_k$ updated
by a discrete feedback controller for a continuous LTI system can be modeled as
a non-deterministic disturbance signal ($\nu_t$ in
Figure~\ref{fig:observersystem}) upper bounded by
%
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, $t_s=T_s-t_d$,
and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm, 
which we extend to sets as $\norm{X}_1 = \sup \{\norm{\vec{x}}_1, \vec{x} \in X\}$.\\ 
For the proof see Appendix~\ref{sec:proofs}
\end{theorem}
%
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
This result allows to over-approximate the effects of the delay by introducing a disturbance signal $\nu_t=[-q_t, q_t]$ at the output.
 %and use the dynamics of the system without delay.  
 This noise will be small when the delay is much smaller than the sampling time ($t_s\approx T_s$ and $\mat{B}_{t_d} \approx 0$).

@@

A safety specification leads to a requirement on the states of the model. 
Consequently, the controller must ensure that the state never violates the
requirement.  In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}^n: = & \forall k \in \{0, \ldots , n\},
  \mat{G}\vec{x}_k \leq \vec{h} \text{ and } \underline{y} \leq y_k \leq \overline{y} \nonumber\\ 
\phi_\mathit{safety}: = & \phi_\mathit{safety}^\infty
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
FWL effects are not explicitly captured here
as they are implicit from Equation \eqref{eq:observer_LTI_FWL_cf}.

@@

The optimizer tool uses infinite precision for the
plant, and may therefore find solutions that the synthesizer cannot, due to
the use of limited precision.  If we cannot optimize the candidate we signal
the {\sc Abstract} phase to use the updated $\phi$ and request {\sc
Synthesize} to produce a new candidate.  Otherwise the optimized candidate
solution is given back to {\sc Verify Synthesis}.

@@

The benchmark suite employed to evaluate the proposed technique is composed
by seven models extracted from the control literature~\cite{acrobot, cstr,
CHEN1979389, KOKOTOVIC198023, gajic2008optimal, Franklin15, maglev,
converters}.  In~order to obtain statistically relevant results, we have
taken these models and modified each possible combination of their
parameters by a 10\% tolerance, thus creating seven hundred total benchmarks.

@@

\begin{table}
\centering
\footnotesize
%
\begin{tabular}{| r | l | c | c | r | r | r | r | r | r | r | r | r |}
%
\hline
\# & \multicolumn{1}{|c|}{Benchmark} & \multicolumn{1}{|c|}{p} & \multicolumn{1}{|c|}{No.} & \multicolumn{4}{|c|}{CEGIS-AR} & \multicolumn{4}{|c|}{CEGIS-OR} \\
   &                                 & & \multicolumn{1}{|c|}{}  & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found} & \multicolumn{1}{|c|}{Min} & \multicolumn{1}{|c|}{Median} & \multicolumn{1}{|c|}{t.o.} & \multicolumn{1}{|c|}{found}\\\hline
1  & DC Motor          & 2 & 100 & 1.3\,s & 26.9\,s& 22& 20& 1.1\,s &   3.8\,s & 2 & 58\\
2  & Helicopter        & 3  & 100 & 57.2\,s & t.o. & 54 & 15 & 7.1\,s  &   18.0\,s & 9 & 90\\
3  & Inverted Pendulum & 4 & 100 &  277\,s& t.o. & 82& 0 & 20.0\,s  &  t.o. & 55 & 20\\
4  & Magnetic Pointer  & 2  & 100 & 5.7\,s & 229.9\,s & 38 & 42 & 3.0\,s  &  9.5\,s & 5 & 95 \\
%5 & Magnetic Suspension          & 2 & 1 & 12.2\,s& 0.6\,s & & & N/A  &   N/A &   N/A\\
5  & Pendulum          & 2 & 100 & 0.8\,s & 53.6\,s & 27 & 71 &0.4\,s  &   4.9\,s & 6 & 92\\
6  & Suspension        & 4 & 100 &15.0\,s & 177.2\,s & 30 & 70 & 14.8\,s&166.9\,s  & 26 & 73\\
7  & Tape Driver       & 3 & 100 & 4.0\,s& 17.7\,s & 11 & 80 &3.1\,s  &   16.8\,s & 8 & 82\\
%9 & Satellite          & 2 & 1 & 9.4\,s& 0.7\,s & & & N/A  &   N/A &   N/A\\
%   & Total                 &   & 2002 & & & & & 1.2\,s & 16.1\,s & 876.8\,s\\
\hline
%
\end{tabular}
\vspace{0.05in}
\caption{\label{tab:cegis_results}
Experimental results.  Minimum, and Median times for standard CEGIS-AR (with
abstraction refinement~\cite{DBLP:conf/cav/AbateBCCDKKP17}) vs CEGIS-OR on
multiple benchmarks.  No.  indicates Number of benchmarks, t.o.~indicates
timeouts (in the median case, it indicates more than half of the benchmarks
timed out).  p denotes the number of variables of a model.}
\end{table}

@@

We found that on average, observer synthesis is over an order of magnitude slower
than full state controller synthesis for the given benchmarks (SAT has
a worst case exponential time complexity, and typical polynomial time, so higher
order systems will show larger differences).
In order to show the improvements due to our optimization-enhanced CEGIS, we
compared our approach against a standard CEGIS with abstraction refinement
procedure~\cite{DBLP:conf/cav/AbateBCCDKKP17} (We note that the naive approach
in~\cite{DBLP:conf/cav/AbateBCCDKKP17} cannot handle the continuous time case.  
Furthermore, since it is in general slower than the abstraction refinement
version, the latter is a more representative benchmark).
For the majority of our benchmarks, we observe that our approach is
significantly faster than~\cite{DBLP:conf/cav/AbateBCCDKKP17}. 
We further note that the number of times CEGIS times out is much larger than
CEGIS-OR which results in more found solutions.
